---
title: "Raincloud Plots"
format:
  html:
    theme: 
      light: flatly
      dark: darkly
    css: styles.css
    toc: true
---


```{r setup, include=FALSE}

# Needed Package names
packages <- c("tidyverse", "kableExtra", "ggplot2", "scales", "RColorBrewer", "zoo", "readxl" , "rmarkdown", "htmltools", "palmerpenguins")
# Install packages not yet installed
installed_packages <- packages %in% rownames(installed.packages())
if (any(installed_packages == FALSE)) {
  install.packages(packages[!installed_packages])
}
# Attach Packages
invisible(lapply(packages, library, character.only = TRUE))

#Plot elements
# gg_text = theme(text = element_text(family = "Times"))
gg_scaley = scale_y_continuous(breaks = pretty_breaks())
gg_scalexd = scale_x_date(breaks = pretty_breaks())

pp <- penguins
```


```{r}
#| code-fold: true
#| code-summary: "Display code"
#| code-link: true
#| code-block-bg: true
#| code-block-border-left: "#31BAE9"
#| code-overflow: wrap
#| layout-ncol: 1



new_labels <- c("Adelie"="ADELIE",
                      "Gentoo"="GENTOO",
                      "Chinstrap"="CHINSTRAP")

rcplot <- function(score, titletext, captext){
  ggplot(pp, aes(x=species, y=bill_length_mm)) + 
    ggdist::stat_halfeye(
      fill=BLUE_50,
      adjust = .5, 
      width = 1.0, 
      .width = 0, 
      justification = -0.3, 
      point_colour = NA,
      alpha=0.6)+
  geom_boxplot(width=0.4, color="black", outlier.shape = NA, alpha=.9, fill=BLUE_50) +
  ggbeeswarm::geom_quasirandom(
    shape=21,                           
    size = 1,
    alpha = 1,
    width = .2,
    bg=BLUE_99,
    color="black") +
  labs(x='', y='', title = titletext, subtitle = captext)+
  # coord_flip()+
    facet_grid(cols=vars(species), scales="free_x", labeller = labeller(species=new_labels))+
  theme_classic()+
  gg_text +
  theme(legend.position="none", legend.title = element_blank(), 
        plot.title = element_text(size=16, face="bold"),
        plot.subtitle = element_text(size=12, lineheight = 0.5 ), axis.text.y=element_text(size=14, face="bold"), axis.text.x=element_blank(),
        axis.ticks.y=element_blank(), axis.title.x=element_text(size=14, face="bold"))

}

plot <- rcplot(scales_all$mspss_score,
              paste0("Bill Length (n=", as.numeric(length(na.omit(pp$bill_length_mm))) , ")"),
                "Bill Length in mm by Species")

plot


